### 2015-07-08 abstract.tex
changes by Tammy

% Even with the space manipulation above, we need to cheat a little
\enlargethispage{2\baselineskip}
Scientific discovery at the exascale will not be possible without significant
new research in the management, storage and retrieval over the long lifespan
of the extreme amounts of data that will be produced.
In this project we will
demonstrate novel techniques to  facilitate efficient mapping of data
objects, even partitioning individual variables, from
the user space onto multiple storage tiers, and enable application-guided
data reductions and transformations to address capacity and bandwidth
bottlenecks. 
Our goal in this project is
to address the associated Input/Output (I/O) and storage challenges in the context of
current and emerging storage landscapes, and expedite insights into mission
critical scientific processes; which is associated with {\bf theme two} of the Funding Opportunity Announcement. To that end, we will build on the
capabilities offered by Adaptable I/O System (ADIOS) and DataSpaces at ORNL that provide I/O abstractions
and services, the Sirocco peer-to-peer file system at
Sandia and the object storage and annotation expertise of UC Santa Cruz, to
explore application and multi-tier storage aware data management
solutions.    This project
brings together a  team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
coordinates these key storage components in a seamless fashion.

Our objective here is to reduce the time to knowledge, not just for a single
application, but for the entire workload in a multi-user environment, where
the storage is shared among users.
We achieve this goal by allowing selectable data quality, by trading its accuracy and error
in order to meet the time or resource constraint. 
We will explore beyond 
checkpoint/restart I/O, and will address the challenges posed by 
key
data access patterns in the knowledge gathering process.
Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
so that the best possible decisions can be autonomically made between the user intentions and
the available system resources.  
We will test our prototypes  on current and future Department of Energy (DOE) system with many of today's applications, including the
XGC1, GTC, QMCPack, and SpecFM3D simulations. 
Our solutions will provide  new functionalities and APIs for
1) specifying, at the application level, data annotations that enable the
quantification of the relative importance and utility of data objects and
enable partitioning  across the storage hierarchy;
2) specifying selectable performance/quality/cost tradeoffs from both the
application and system perspectives;
3) evaluating these tradeoffs at runtime during data placement and movement,
and executing the resultant policies in an autonomic system using models,
heuristics and continuous learning; and
4) leveraging techniques such as application-aware data compression, 
and I/O prioritization to enforce these
policies.
\noindent

>>>>

% Even with the space manipulation above, we need to cheat a little
\enlargethispage{2\baselineskip}
Scientific discovery at the exascale will not be possible without significant
new research in the management, storage and retrieval over the long lifespan
of the extreme amounts of data that will be produced.
In this project we will
demonstrate novel techniques to  facilitate efficient mapping of data
objects, even partitioning individual variables, from
the user space onto multiple storage tiers, and enable application-guided
data reductions and transformations to address capacity and bandwidth
bottlenecks. 
Our goal in this project is
to address the associated Input/Output (I/O) and storage challenges in the context of
current and emerging storage landscapes, and expedite insights into mission
critical scientific processes; which is associated with {\bf theme two} of the Funding Opportunity Announcement. To that end, we will build on the
capabilities offered by Adaptable I/O System (ADIOS) and DataSpaces at ORNL that provide I/O abstractions
and services, the Sirocco peer-to-peer file system at
Sandia, and the object storage and annotation expertise of UC Santa Cruz, to
explore application and multi-tier storage aware data management
solutions.    This project
brings together a  team with strong expertise in I/O middleware (ORNL, Rutgers), file system (SNL, UCSC) and storage (UCSC), and connects and
coordinates these key storage components in a seamless fashion.

Our objective here is to reduce the time to knowledge, not just for a single
application, but for the entire workload in a multi-user environment, where
the storage is shared among users.
We achieve this goal by allowing selectable data quality, by trading its accuracy and error
in order to meet the time or resource constraint. 
We will explore beyond 
checkpoint/restart I/O, and will address the challenges posed by 
key
data access patterns in the knowledge gathering process.
Ultimately, we will take the knowledge from the storage system to provide vital feedback to the middleware 
so that the best possible decisions can be autonomically made between the user intentions and
the available system resources.  
We will test our prototypes  on current and future Department of Energy (DOE) system with many of today's applications, including the
XGC1, GTC, QMCPack, and SpecFM3D simulations. 
Our solutions will provide  new functionalities and Application Program Interfaces (API) for:
1) specifying, at the application level, data annotations that enable the
quantification of the relative importance and utility of data objects and
enable partitioning  across the storage hierarchy;
2) specifying selectable performance/quality/cost tradeoffs from both the
application and system perspectives;
3) evaluating these tradeoffs at runtime during data placement and movement,
and executing the resultant policies in an autonomic system using models,
heuristics and continuous learning; and
4) leveraging techniques such as application-aware data compression, 
and I/O prioritization to enforce these
policies.
\noindent

### 2015-07-08 coverpage.tex
changes by Tammy

%% Lead PI name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Lead Principal Investigator:}} \=
\vrule height0.4pt depth 0pt width 2.5in \\
\> \> Scott Klasky,  Distinguished Scientist \\
\> \> Oak Ridge National Laboratory \\
\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, Tennessee 37831-6057 \\
\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
\> \> Email: klasky@ornl.gov
\end{tabbing}

@@

%% Adminisrative Point of Contact name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Administrative Point of Contact:}} \=
\vrule height0.4pt depth 0pt width 2.5in \\
\> \> Barney Maccabe \\
\> \> Oak Ridge National Laboratory \\
\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, TN 37831-6057 \\
\> \> (T) 865-241-6504 \\
\> \> Email: maccabe@ornl.gov
\end{tabbing}

>>>>

%% Lead PI name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Lead Principal Investigator:}} \= Scott Klasky,  Distinguished Scientist \\
\> \> Oak Ridge National Laboratory \\
\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, Tennessee 37831-6057 \\
\> \> (T) 865-241-9980 (Fax) 865-241-4811 \\
\> \> Email: klasky@ornl.gov
\end{tabbing}
%\vrule height0.4pt depth 0pt width 2.5in \\

@@

%% Adminisrative Point of Contact name, telephone number, email:
\begin{tabbing}
\hspace*{10mm} \= \makebox[65mm][l]{\textbf{Administrative Point of Contact:}} \= Barney Maccabe \\
\> \> Oak Ridge National Laboratory \\
\> \> Computer Science and Mathematics Division \\
\> \> Oak Ridge, TN 37831-6057 \\
\> \> (T) 865-241-6504 \\
\> \> Email: maccabe@ornl.gov
\end{tabbing}
%\vrule height0.4pt depth 0pt width 2.5in \\

### 2015-07-08 narrative/data-description.tex
changes by Tammy

\paragraph{Proposed research approach:} 
We will design and develop new techniques for data management, description,
and access that allow users to describe the data utility based on their
expectations. This can be done by allowing users to apply plugins
provide data-specific functionality to the middleware and storage system.
These plugins may be executed by the system to
calculate the utility value for each chunk, handle custom compression
and decompression, and perform other data-specific services, thus allowing
the system to more effectively manage data placement and retrieval.
%
%We will design an updated I/O API that will expand on the current POSIX I/O
%semantics by adding new information to the I/O request.

@@

We will also develop a new type of querying system that allows an application
to get estimated timing information for operations on the storage system.
We will explore
how these query functions can be integrated into common applications with
minimal code disruptions and the set of policies that applications can use to
respond to this new information. We will also explore how applications can
query the storage system on space availability across different tiers and how
applications their tailor their output data to make the most effective use of
available space.
Finally we will study the use of an external data annotation system that can
provide information to the storage system without requiring recompilation of the application.

%
From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
the system, 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
Sec~\ref{sec:data-refactor}) which will classify the particle data into 4 different
prioritization levels. The user can then pick a data utility, such as 3 months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest priority element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
%

@@

Finally we will investigate the use of additional information
on top of the data model.
This information will be used to indicate data associations that exceed
traditional data models. For example, the fields and mesh in our example have
an inherent relationship. A common data model will describe all of the
variables in the mesh (the coordinates, the units, and the types of elements if
this were a finite-element mesh). We will explore new mechanisms that will
group data together making it explicit that M1, F1, and G1 need to be organized
together.  In other words, it will not be helpful if data from M1 and M2 were
organized together along with only F1. When the users wants to visualize data
from the field, F1, they will need only M1, and M2 would be additional
information. Likewise, if the user wants to see F1 and F2, then they will need
the mesh from M1 and M2. This requires that the time to access all of this
data is approximately the same.
We will investigate what choices the SSIO system must make in order to
keep these pieces ``together'', from a performance perspective. Can the storage
system move everything to a lower layer when all of it doesn't fit or is there
other, less tightly related data that can be evicted to preserve the collective
retrieval performance?  We need to investigate the potential problems when this
occurs and how the users can specify additional information to ensure they get
the best possible data from the SSIO layer.

>>>>

\paragraph{Proposed research approach:} 
We will design and develop new techniques for data management, description,
and access that allows users to describe the data utility based on their
expectations. This can be done by allowing users to apply plugins
provide data-specific functionality to the middleware and storage system.
These plugins may be executed by the system to
calculate the utility value for each chunk, handle custom compression
and decompression, and perform other data-specific services, thus allowing
the system to more effectively manage data placement and retrieval.
%
%We will design an updated I/O API that will expand on the current POSIX I/O
%semantics by adding new information to the I/O request.

@@

We will also develop a new type of querying system that allows an application
to get estimated timing information for operations on the storage system.
We will explore
how these query functions can be integrated into common applications with
minimal code disruptions and the set of policies that applications can use to
respond to this new information. We will also explore how applications can
query the storage system on space availability across different tiers and how
applications their tailor their output data to make the most effective use of
available space.
Finally, we will study the use of an external data annotation system that can
provide information to the storage system without requiring recompilation of the application.

%
From an application level, we will explore these techniques through our ADIOS
framework using many of our existing application partners, including
XGC1, GTC, and SPECFM3D. We will investigate: 1) what hints the users can give
the system, and 2) how the users can replace much of their
application specific code with
``system level code'' through the use of these hints.  We will design an API
that will be integrated with the storage layer to give time estimations so that
users can decide which data priorities will be written and read. The utility
value for each chunk will be used to sort the chunks into prioritized bins.
Chunks in a particular bin will be assigned an appropriate storage
tier and lifetime, which can be specified by the user through the API.
As an example, Figure~\ref{fig:ssio-bucket} illustrates a use case for the XGC1 code.
The user can pick a data-refactoring scheme (see
Sec~\ref{sec:data-refactor}) which will classify the particle data into four different
prioritization levels. The user can then pick a data utility, such as three months
for the highest priority items and 12 months for the lowest priority data.
The user will then call an API function that allows them
to specify all of the chunks they want
written (particles, two fields, and a mesh in this example) and the system
will return a time estimate for this write operation.
If this time estimate is too large, an API call can be made to 
specify the lowest priority element to be written, skipping the rest in order to save time.
The data will then be stored
across the layers of the file system where the two highest priority buckets
will be stored in the Parallel File System, but the next highest priority items
will be stored lower in the storage stack. Depending on the capabilities of the actual storage system, 
SSIO will decide if the data has to go through the parallel file system 
towards a lower stack or can be staged to that stack directly. 
%

@@

Finally we will investigate the use of additional information
on top of the data model.
This information will be used to indicate data associations that exceed
traditional data models. For example, the fields and mesh in our example have
an inherent relationship. A common data model will describe all of the
variables in the mesh (the coordinates, the units, and the types of elements if
this were a finite-element mesh). We will explore new mechanisms that will
group data together making it explicit that M1, F1, and G1 need to be organized
together.  In other words, it will not be helpful if data from M1 and M2 were
organized together along with only F1. When the users wants to visualize data
from the field, F1, they will need only M1, and M2 would be additional
information. Likewise, if the user wants to see F1 and F2, then they will need
the mesh from M1 and M2. This requires that the time to access all of this
data is approximately the same.
We will investigate what choices the SSIO system must make in order to
keep these pieces "together,'' from a performance perspective. Can the storage
system move everything to a lower layer when all of it doesn't fit, or is there
other, less tightly related data that can be evicted to preserve the collective
retrieval performance?  We need to investigate the potential problems when this
occurs and how the users can specify additional information to ensure they get
the best possible data from the SSIO layer.

### 2015-07-08 narrative/data-reduction.tex
changes by Tammy

The classical workflow where the entire dataset is written to storage for
later analysis is no longer viable simply because the amount of generated data is too large.
In future, it will be vital to take advantage of a priori
information when 1) writing and reading from the SSIO layer in order to gain
performance and predictability, and to prioritize I/O so that the data that
is most useful for end-user can be written in the time available, and 2)
performing {\em in situ} operations and analysis before storing the information.

@@

\paragraph{State of the art:} {\bf Data Reduction Techniques} Data compression
broadly consists of three basic steps: (1) Pre-processing to extract as much
structure as feasible. This step is variously referred to as filtering,
preconditioning, and preprocessing. But in all cases, the objective is to
represent the data in a form that becomes more amenable to compression.  (2)
Removal of redundancy such as omission of duplicates etc.  Finally, (3) the
resulting data is compressed using entropy coding. The level of compression
that can be achieved in the final step is limited by the Shannon information
content of the data \cite{Gray:book}, and many standard tools are available for
efficient and perfect entropy coding. The potential for compression through
removal of redundancy is generally minor. The main scope for achieving
significant compression (e.g. a compression rate of two) therefore
lies in the pre-processing step.

Research in the construction of pre-processing techniques is still in its
infancy, especially in context of exascale simulations. Techniques that have
been advocated range from simple difference filtering, bit-sorting, prefix
transformation through to simply reordering the data.  What these, and other
catch-all methods, fail to exploit is a priori information on the nature and
provenance of the data.

@@

\item Precision based re-organization. This is where the most significant bytes
	of the data are all grouped together from each object. This data
	generally will have a higher utility then the data with the least
	significant bytes.  This process has the following three steps which
	need to be efficiently implemented when data gets placed to the storage
	system and from the storage system.  First, the data needs to be
	re-arranged which involves memory operations and needs to be done
	in-situ. This operation involves no communication.  Data needs to be
	copied, so the memory requirements are increased. At most this requires
	two copies an individual dataset in memory, and we will investigate
	techniques to allocate and deallocate this memory if the user will
	specify that the data will be overwritten after it is written to the
	storage system. This is often the case for many of the data quantities
	written from the simulation, but there are many cases where, for
	example, we want to write all of the particles from a Particle In Cell
	(PIC) simulation.  Since the particles will be used later in the
	calculation, we need to duplicate the storage. Our observation with
	working with the XGC1, GTC, Warp, and PiconGPU simulations (all
	leadership class simulations) is that we can temporary increase the
	storage of the particles, and then release them since the temporary
	arrays used in the calculations are often freed when a PIC iteration is
	finished. 
%
\item Frequency-based re-organization.  Another common technique to classify
	the importance of data is to first re-organize the data in frequency
	space and then specify that as the frequency increases the importance
	of data decreases.   Common technique which is used for streaming data,
	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to
	code streams which have regions of interest that offer several
	mechanisms to support spatial random access or region of interest
	access at varying degrees of granularity. It is possible to store
	different parts of the same data using different quality.  This allows
	us to place the lowest frequency pieces in the fastest storage, and the
	highest frequency either on the slowest storage tiers, or if the data
	sizes are prohibitively  costly, not even write out the highest
	frequency. This binning of data in frequency space can take as an input
	(storage bandwidth knowledge, storage size limitations, and user
	intentions of how long the majority of the data will be frequently
	accessed.  In order to fully take advantage of frequency based
	re-organization we have devised a scheme where the data pre-conditioner
	first sorts the data in bins of 1024 words, and then we maintain the
	sort order, and then we used wavelets and spline fits to then
	re-organize the values.  This re-organization allows us to understand
	that the resultant data is smoother now, and can be reduced in each one
	of the bins.  \end{itemize}

@@

We will also research the time it takes to re-organize the data and use the
multi-layer approach to staging to understand where the data gets reduced and
where it gets re-organized. Open questions are: 1) do we move the data over to
other nodes to then reduce and re-organize, 2) do we first reduce the data and
then later re-organize this data on different nodes?  3) do we use different
cores on a node to do data-refactoring.  Similarly we will examine the cost of
reading the data. It is clear that if we only read in the high-importance data
it should improve the overall time to move the data from the SSIO layer to the
application, but what happens when users want all of the data? There is an
obvious penalty of having to take data in a hierarchical layout and then place
this back to a single-layer view.  Additional research questions which must be
addressed: Will some types of data (e.g. PIC data) be more
conducive  for data-refactoring than others? How much semantic-domain knowledge
will we need to get a dramatic improvement of the storage system over not doing
anything special?

>>>>

The classical workflow where the entire dataset is written to storage for
later analysis is no longer viable simply because the amount of generated data is too large.
In the future, it will be vital to take advantage of a priori
information when 1) writing and reading from the SSIO layer in order to gain
performance and predictability, and to prioritize I/O so that the data that
is most useful for end-user can be written in the time available, and 2)
performing {\em in situ} operations and analysis before storing the information.

@@

\paragraph{State of the art:} {\bf Data Reduction Techniques.} Data compression
broadly consists of three basic steps: (1) Pre-processing to extract as much
structure as feasible. This step is variously referred to as filtering,
preconditioning, and preprocessing. But in all cases, the objective is to
represent the data in a form that becomes more amenable to compression.  (2)
Removal of redundancy such as omission of duplicates, etc.  Finally, (3) the
resulting data is compressed using entropy coding. The level of compression
that can be achieved in the final step is limited by the Shannon information
content of the data \cite{Gray:book}, and many standard tools are available for
efficient and perfect entropy coding. The potential for compression through
removal of redundancy is generally minor. The main scope for achieving
significant compression (e.g. a compression rate of two) therefore
lies in the pre-processing step.

Research in the construction of pre-processing techniques is still in its
infancy, especially in context of exascale simulations. Techniques that have
been advocated range from simple difference filtering, bit-sorting, prefix
transformation through to simply reordering the data.  What these, and other
catch-all methods fail to exploit is a priori information on the nature and
provenance of the data.

@@

\item Precision based re-organization. This is where the most significant bytes
	of the data are all grouped together from each object. This data
	generally will have a higher utility then the data with the least
	significant bytes.  This process has the following three steps which
	need to be efficiently implemented when data gets placed to the storage
	system and from the storage system.  First, the data needs to be
	re-arranged which involves memory operations and needs to be done
	in-situ. This operation involves no communication.  Data needs to be
	copied, so the memory requirements are increased. At most this requires
	two copies an individual dataset in memory, and we will investigate
	techniques to allocate and deallocate this memory if the user will
	specify that the data will be overwritten after it is written to the
	storage system. This is often the case for many of the data quantities
	written from the simulation, but there are many cases where, for
	example, we want to write all of the particles from a Particle In Cell
	(PIC) simulation.  Since the particles will be used later in the
	calculation, we need to duplicate the storage. Our observation with
	working with the XGC1, GTC, Warp, and PiconGPU simulations (all
	leadership class simulations) is that we can temporary increase the
	storage of the particles, and then release them since the temporary
	arrays used in the calculations are often freed when a PIC iteration is
	finished. 
%
\item Frequency-based re-organization.  Another common technique to classify
	the importance of data is to first re-organize the data in frequency
	space and then specify that as the frequency increases the importance
	of data decreases.   Common technique which is used for streaming data,
	and data reduction techniques such as JPEG-2000~\cite{jpeg2000} is to
	code streams which have regions of interest that offer several
	mechanisms to support spatial random access or region of interest
	access at varying degrees of granularity. It is possible to store
	different parts of the same data using different quality.  This allows
	us to place the lowest frequency pieces in the fastest storage, and the
	highest frequency either on the slowest storage tiers, or if the data
	sizes are prohibitively  costly, not even write out the highest
	frequency. This binning of data in frequency space can take as an input
	(storage bandwidth knowledge, storage size limitations, and user
	intentions of how long the majority of the data will be frequently
	accessed.  In order to fully take advantage of frequency based
	re-organization we have devised a scheme where the data pre-conditioner
	first sorts the data in bins of 1024 words, then we maintain the
	sort order, and then we used wavelets and spline fits to then
	re-organize the values.  This re-organization allows us to understand
	that the resultant data is smoother now, and can be reduced in each one
	of the bins.  \end{itemize}

@@

We will also research the time it takes to re-organize the data and use the
multi-layer approach to staging to understand where the data gets reduced and
where it gets re-organized. Open questions are: 1) do we move the data over to
other nodes to then reduce and re-organize, 2) do we first reduce the data and
then later re-organize this data on different nodes?, or 3) do we use different
cores on a node to do data-refactoring.  Similarly we will examine the cost of
reading the data. It is clear that if we only read in the high-importance data
it should improve the overall time to move the data from the SSIO layer to the
application, but what happens when users want all of the data? There is an
obvious penalty of having to take data in a hierarchical layout and then place
this back to a single-layer view.  Additional research questions which must be
addressed: Will some types of data (e.g. PIC data) be more
conducive  for data-refactoring than others? How much semantic-domain knowledge
will we need to get a dramatic improvement of the storage system over not doing
anything special?

### 2015-07-08 narrative/intro.tex
changes by Tammy

% 
Our proposed research program is guided by the following overarching objectives:
\begin{itemize}
\item
{\em Address emerging storage architecture and technology landscape:}
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes,
% and expedite insights into critical
%scientific processes, while demonstrating the validity of our approach in key
%DOE domains. 
%We will research techniques for creating and using a Scalable 
%Storage Software Infrastructure, and integrating services from the middleware 
%layer with Storage System capabilities to support 
and will develop novel strategies for distributing 
data both horizontally and vertically across the layers of these storage systems.  
Negotiation between these layers will be provided as a fundamental service, allowing users to
easily save ``the best''  data, which is automatically placed throughout the
entire storage stack.
\item
{\em Develop a science driven, system aware approach:}
%Our approach targets the expected characteristics of exascale storage hardware.
%The storage layer will be partitioned into multiple heterogeneous tiers with
%vastly different performance characteristics. The layers will be further
%differentiated by the constraints on capacity and data lifetime within each
%layer. Tape archival storage will still maintain data long term, but access to
%this data will be orders of magnitude slower than the next layer and naively
%accessing data from archival storage will greatly impact productivity.
%
Given that storage systems will be composed of multiple heterogeneous tiers with
vastly different performance characteristics and constraints on capacity and 
supported data lifetime, our approach will enable researchers to 
place more knowledge into the SSIO system so that data can be recast from 
a sequence of bytes to information rich data. This information rich data can then
be used by the system to guide the  placement within the tiered storage and will optimize the time
to gain knowledge and insight from exascale simulations.

@@

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of IO middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of IO middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
\hline
\end{tabular}
\end{center}
%\caption {WRITE A CAPTION FOR THIS TABLE}
\vskip -0.5cm
\caption {}
\label{tab:call}
\end{table}

>>>>

% 
Our proposed research program is guided by the following overarching objectives:
\begin{itemize}
\item
{\em Address emerging storage architecture and technology landscape:}
We will address the associated I/O and storage challenges in the context of
current and emerging storage landscapes,
% and expedite insights into critical
%scientific processes, while demonstrating the validity of our approach in key
%DOE domains. 
%We will research techniques for creating and using a Scalable 
%Storage Software Infrastructure, and integrating services from the middleware 
%layer with Storage System capabilities to support 
and will develop novel strategies for distributing 
data both horizontally and vertically across the layers of these storage systems.  
Negotiation between these layers will be provided as a fundamental service, allowing users to
easily save ``the best''  data, which is automatically placed throughout the
entire storage stack.
\item
{\em Develop a science driven, system aware approach:}
%Our approach targets the expected characteristics of exascale storage hardware.
%The storage layer will be partitioned into multiple heterogeneous tiers with
%vastly different performance characteristics. The layers will be further
%differentiated by the constraints on capacity and data lifetime within each
%layer. Tape archival storage will still maintain data long term, but access to
%this data will be orders of magnitude slower than the next layer and naively
%accessing data from archival storage will greatly impact productivity.
%
Given that storage systems will be composed of multiple heterogeneous tiers with
vastly different performance characteristics and constraints on capacity and 
supported data lifetime, our approach will enable researchers to 
place more knowledge into the Storage Systems and Input/Output (SSIO) system so that data can be recast from 
a sequence of bytes to information rich data. This information rich data can then
be used by the system to guide the  placement within the tiered storage and will optimize the time
to gain knowledge and insight from exascale simulations.

@@

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf See} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations for users to guide the system for placement  & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Create the ability to re-organize, reduce and prioritize data chunks from an object& \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & what are we doing. & \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of I/O middleware with the resource management system & We will manage data from the application  to the middleware to the storage system and allow them to have an autonomic interaction & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & We will create APIs to guide the placement of chunks of data across the SSIO and autonomically move data when necessary& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 & Approaches to improve the ability of SSIO software to support checkpoint/restart & We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then purged when it is no longer needed & \S \ref{fig:ssio-bucket}  \\
\hline
\end{tabular}
\end{center}
%\caption {WRITE A CAPTION FOR THIS TABLE}
\vskip -0.5cm
\caption {}
\label{tab:call}
\end{table}

### 2015-07-08 narrative/placement.tex
changes by Tammy

Our proposed research will address data management challenges across the entire 
data life cycle in an application driven and system aware manner. 
%given our thesis that more knowledge, and more specific knowledge about the
%data, can allow the management of data to be optimized with reduced cost.
Some of the high level research questions that we will explore are as follows: 
(1) How can we initially place data so that it can be discovered and consumed efficiently?
%What methods can be utilized in the initial placement of data that do not impose expensive
%costs on discovery, consumption and migration? 
(2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
optimized at runtime, both from the application and system perspective?
%(3) How can we migrate data throughout the layers of storage
%without making the discovery process unbounded? 
(4) How can knowledge about the application used to better prepare the data for consumption? 
(5) When and how do we make the decision to purge data?  %using this additional knowledge?

>>>>

Our proposed research will address data management challenges across the entire 
data life cycle in an application driven and system aware manner. 
%given our thesis that more knowledge, and more specific knowledge about the
%data, can allow the management of data to be optimized with reduced cost.
Some of the high level research questions that we will explore are as follows: 
(1) How can we initially place data so that it can be discovered and consumed efficiently?,
%What methods can be utilized in the initial placement of data that do not impose expensive
%costs on discovery, consumption and migration? 
(2) How can the placement and migration of data across a multi-tiered storage hierarchy be 
optimized at runtime, both from the application and system perspective?,
%(3) How can we migrate data throughout the layers of storage
%without making the discovery process unbounded? 
(4) How can knowledge about the application used to better prepare the data for consumption?, and 
(5) When and how do we make the decision to purge data?  %using this additional knowledge?
